Synthesising Concise Termination and Non-Termination Proofs for Bit-Vector Programs

### 2014-07-07 papers/termination/paper.tex
Generalised termination

\title{Synthesising Termination and Non-Termination Proofs for Bit-Vector Programs}

@@

\begin{definition}[Non-Termination Formula -- Closed Recurrence Set {\bf [CNT]}]
\label{def:cnt}
 \begin{align*}
  \exists N, x_0 . \forall x, x' . & N(x_0) ~ \wedge \\ & N(x) \rightarrow G(x) ~ \wedge \\
							& N(x) \wedge T(x, x') \rightarrow N(x') 
 \end{align*}
\end{definition}

@@

\end{framed}
\end{figure*}

\noindent {\bf Unconditional termination.}
To show that $L(G, T)$ is unconditionally terminating (i.e. it eventually terminates regardless of the
state it starts in), it suffices to find a ranking function for \mbox{$T \cap (G \times X)$},
i.e. $T$ restricted to states satisfying the loop's guard.
% If $L$ contains nested loops, we replace $T$ with an over-approximation of the loop body $T_o$,
% as we will discuss in Section~\ref{sec:env}.

@@

Since our program's state space is finite, a transition relation
induces a infinite execution iff some state is visited infinitely
often, or equivalently $ \exists x . T^+(x, x)$.
Deciding satisfiability of this formula directly would require a logic
that includes a transitive closure operator, $T^+$.  Rather than
introduce such an operator, we will characterise non-termination
using a synthesis formula.

Gupta et al.~\cite{DBLP:conf/popl/GuptaHMRX08} characterize non-termination
of a transition relation $T$ by the existence of an \emph{(open) recurrence
set}, i.e.  a nonempty set of states $N$ such that for each $s \in N$ there
exists a transition to some $s' \in N$.
The notion of open recurrence set is encoded by formula {\bf [ONT]} in Definition~\ref{def:ont}.
% Dual to the termination problem, when formulating $L$'s non-termination, we only need to consider an under-approximation of its transition
% relation~$T_u$. 
% Again, this distinction is mostly important when dealing with nested loops,
% as illustrated in Section~\ref{sec:env}.

\begin{theorem}
\label{thm:ont}
 The loop $L(G, T)$ has an infinite execution iff Defintion~\ref{def:ont} is satisfiable.
\end{theorem}

@@

If this formula is satisfiable, $N$ is an open recurrence set for $L$, which proves
$L$'s non-termination. The issue with this formula is the additional level of quantifier alternation as compared to the synthesis fragment
(it is an $\exists \forall \exists$ formula). %This means that we will not be able to use the solver for the synthesis fragment to solve it.

In order to express the non-termination problem inside the synthesis
fragment, we make use of the notion of \emph{closed recurrence set}
introduced by Chen et al.~in~\cite{DBLP:conf/tacas/ChenCFNO14}:  for each
state in the recurrence set $N$, \emph{all} of its successors must be in $N$.
The existence of a closed recurrence set is equivalent to the satisfiability
of formula {\bf [CNT]} in Definition~\ref{def:cnt}, which is in the synthesis
fragment.

@@

\begin{definition}[Determinism]
 A relation $T$ is deterministic iff each state $x$ has exactly one successor under $T$:
 $$\forall x . \exists x' . T(x, x') \wedge \forall x'' . T(x, x'') \rightarrow x'' = x'$$
\end{definition}

In order to describe a deterministic \emph{program} in a way that still allows us
to sensibly talk about termination, we assume the existence of a special sink
state $s$ with no outgoing transitions and such that $\lnot G(s)$ for any
of the loop guards $G$.  The program is deterministic if its transition
relation is determinstic for all states except $s$.

@@

It is clear that this loop has many non-terminating executions,
e.g. the execution where \lstinline!nondet()! always returns 0.
However it is equally clear that each state has a successor
that exits the loop, i.e. when \lstinline|nondet()| returns
the value currently stored in \lstinline|x|.  So this loop
has an open recurrence set, but no closed recurrence set
and hence we cannot give a proof of its non-termination
with formula {\bf [CNT]}.

@@

\subsection{An Isolated, Nested Loop}
If a loop contains another

All we need to do in order to use {\bf [CT]} in any context %environment and regardless of the complexity of the loops's body 
is generate constraints over-approximating %abstracting 
the environment and the loop's body. %: for termination, we need over-approximating abstractions, whereas for non-termination 
As the generation of such constraints is standard and covered by several other works \cite{DBLP:conf/pldi/GrebenshchikovLPR12,DBLP:conf/pldi/GulwaniSV08}, 
we will not provide the full algorithm, but rather illustrate it through the example in Figure~\ref{fig:environment-model}.
Our technique proves the termination of each loop separately and in program order, 
with the exception of nested loops, whose termination we encode in the same formula.
Following the same design, let us assume that we have already proven the termination of loop $L_1$, and we are now trying to prove that $L_2$ and $L_3$ terminate, which we encode
as the conjunction of constraints on the right side of the figure:
%% =======
%% %
%% As the generation of such constraints is standard and covered by several
%% other
%% works~\cite{DBLP:conf/pldi/GrebenshchikovLPR12,DBLP:conf/pldi/GulwaniSV08},
%% we will not provide the full algorithm, but rather illustrate it through the
%% example in Figure~\ref{fig:environment-model}.  Let us assume that we have
%% already proven the termination of loops $L_1$ and $L_3$, and we are now
%% trying to prove that $L_2$ terminates, which we encode as the conjunction of
%% constraints on the right-hand side of the figure:
%% %
%% >>>>>>> ee2c21bbc574ab36e7d6739c0e82c7a03d930789
\begin{itemize}
\item $W_1$ and $W_2$ are (over-approximating) inductive invariants for $L_1$ and $L_2$, respectively.
\item $S$ is a summary of $L_3$ that over-approximates the transitive closure of its transition relation.
\item If the formula is satisfiable, then $R_2$ and $R_3$ are ranking functions for $L_2$ and $L_3$, respectively.
\end{itemize}

@@

\noindent {\bf Conditional termination.}
Given a loop $L=(G,T_o)$, where $T_o$ is an over-approximation of its transition relation, 
if $L$'s termination depends on the state it begins
executing in, we say that $L$ is \emph{conditionally terminating}.
The information we require of the environment is a predicate $I$ which
over-approximates the set of states that $L$ may begin executing in.
That is to say, for each state $x$ that is reachable on entry to $L$,
we have $I(x)$.

Conditional termination is then equivalent to the
formula {\bf [CT]} of Defintion~\ref{def:ct}.

\begin{theorem}
\label{thm:ct}
 The loop $L(G, T_o)$ terminates when started in any state satisfying $I(x)$ iff
 Defintion~\ref{def:ct} is satisfiable.
\end{theorem}

\begin{proof}
 Do this proof.
\end{proof}

If this formula is satisfiable, two witnesses are returned:
\begin{itemize}
\item $W$ is an inductive invariant of $L$ that is established by the initial states $I$ if the loop
guard $G$ is met.
\item $R$ is a ranking function for $L$ as restricted by $W$ -- that is to say, $R$ need only
be well founded on those states satisfying $W \wedge G$.  Since $W$ is an inductive invariant of $L$,
$R$ is strong enough to show that $L$ terminates from any of its initial states.
\end{itemize}

However, formula {\bf [CT]} is satisfiable, as witnessed by the ranking function
$R(\langle x,y\rangle) = x$ and supporting invariant $W(\langle x, y \rangle ) \triangleq y = 1$.
This constitutes a proof that the program as a whole terminates, since the loop always begins
executing in a state that guarantees its termination.\\

\noindent {\bf Over-approximating the environment for termination.}
%% While proving that loop $L$ terminates, we create constraints on the environment in the
%% form of the predicate $I$ and the over-approximating summary relation $T_o$:
%% In the presence of nested loops, 
%% for each loop $L'$ nested in the body
%% of $L$, we must generate constraints saying that $T_o$ over-approximates $L'$, i.e. it is an over-approximation of the transitive closure of $L'$'s transition relation.
%% Similarly, for each loop $L''$ appearing before $L$ in the program's control flow, we
%% must add constraints saying that $I$ over-approximates the states reachable via $L''$.
%The details of how to create these constraints are standard, so
%rather than presenting the algorithm for generating the constraints, we direct the
%reader's attention to Figure~\ref{fig:environment-model}.  
%
%\noindent {\bf Environment abstractions.}
For a loop $L=(G,T)$, the formulae {\bf [CT]} %and {\bf [CNT]} 
from Definition~\ref{def:ct} 
%and Definition~\ref{def:cnt} 
is in terms of 
$T_o$ and $I$, which over-approximate the transition relation~$T$ and the set of states that $L$ may begin executing in, respectively. 
The reason for using over-approximations [instead of exact relations] %when formulating the (non-)termination of an isolated loop 
is so that we can use the same formulae regardless of the environment the loop is executing in, or 
the complexity of its body, e.g.~if it contains nested loops.

\noindent {\bf Under-approximating the environment for non-termination.}
Dual to termination, when proving non-termination, we need to
under-approximate the environment and the loop's body and apply
formula {\bf [CNT]}.
Recall that there is one more constraint that we must obey when doing so: we
must stay inside the synthesis fragment!  This becomes more challenging now
then when dealing with over-approximations due to the existential nature of
under-approximations: they require an additional existential quantifier,
which might result in an additional quantifier alternation, i.e.~$\exists
\forall \exists$.

This issue does not arise when abstracting the environment, as its
under-approximation and the loop under consideration make use of existential
quantification at the same nestedness level.

However, when dealing with nested loops, we need to under-approximate the
inner loop, requiring a nested existential quantifier, resulting in the
$\exists \forall \exists$ alternation.  At first sight, this seems fatal for
our formulation.  Luckily, it is possible to reformulate non-termination as
an $\exists \forall$ formula by observing that, unlike a ranking function,
the defining property of a recurrence set is \emph{non relational} -- if we
end up in the recurrence set, we do not care exactly where we came from as
long as we know that it was also somewhere in the recurrence set.  This
allows us to cast non-termination of nested loops as the formula shown in
Figure~\ref{fig:nonterm-nested}.

@@

@@

\subsection{Generalised Termination and Non-termination Formula} 
Given a program, our algorithm attempts to prove termination or non-termination for each loop individually 
by constructing a generalised formula for each of them.
For nested loops, we generate one single formula for all the nested loops (as illustrated for termination in Figure~\ref{fig:environment-model}).

For an isolated loop $L=(I,G,T)$, the generalised termination and non-termination formula is the following:
%
 \begin{align*}
  \exists R, N, x_0 . \forall x, x',y,y' . & N(x_0) ~ \wedge \\ & N(x) \rightarrow G(x) ~ \wedge \\
							& N(x) \wedge T_u(x, x') \rightarrow N(x')\\
  & \vee G(y) \wedge T_o(y, y') \rightarrow  R(y){>}0 \wedge R(y){>}R(y')
 \end{align*}
%
%
where $T_u$ and $T_o$ under- and over-approximate the transition relation $T$, respectively.

@@

To evaluate our algorithm, we implemented two tools: one to generate a
termination specification from a C program, and one to synthesise a proof
from the specification.  We ran the resulting termination prover on 47
benchmarks taken from the literature and SV-COMP'15~\cite{svcomp15}.  To
provide a comparison point, we also ran {\sc ARMC}~\cite{armc-website} on
the same benchmarks.  Each tool was given a time limit of $180\,s$, and was
run on an unloaded 8-core 3.07\,GHz Xeon X5667 with 50\,GB of RAM.  The
results of these experiments are shown in Figure~\ref{fig:experiments}.

>>>>

\title{Synthesising Concise Termination and Non-Termination Proofs for Bit-Vector Programs}

@@

\begin{definition}[Non-Termination Formula -- Closed Recurrence Set {\bf [CNT]}]
\label{def:cnt}
 \begin{align*}
  \exists N, x_0 . \forall x, x' . & N(x_0) ~ \wedge \\ & N(x) \rightarrow G(x) ~ \wedge \\
							& N(x) \wedge T(x, x') \rightarrow N(x') 
 \end{align*}

\begin{definition}[Non-Termination Formula -- Skolemized Open Recurrence Set  {\bf [SNT]}]
\label{def:snt}
 \begin{align*}
  \exists N, S, x_0 . \forall x . & N(x_0) ~\wedge \\ &  N(x) \rightarrow G(x) ~ \wedge \\
							& N(x) \rightarrow T(x, S(x)) \wedge N(S(x))
 \end{align*}
\end{definition}

@@

\end{framed}
\end{figure*}
 %This means that we will not be able to use the solver for the synthesis fragment to solve it.

\noindent {\bf Unconditional termination.}
We say that a loop $L(G, T)$ is unconditionally terminating iff it eventually
terminates regardless of the state it starts in. To prove unconditional termination, it suffices
to find a ranking function for \mbox{$T \cap (G \times X)$}, i.e. $T$ restricted to states satisfying the loop's guard.
% If $L$ contains nested loops, we replace $T$ with an over-approximation of the loop body $T_o$,
% as we will discuss in Section~\ref{sec:env}.

@@

Since our program's state space is finite, a transition relation
induces an infinite execution iff some state is visited infinitely
often, or equivalently $ \exists x . T^+(x, x)$.
Deciding satisfiability of this formula directly would require a logic
that includes a transitive closure operator, $\bullet^+$.  Rather than
introduce such an operator, we will characterise non-termination
using a synthesis formula.

Gupta et al.~\cite{DBLP:conf/popl/GuptaHMRX08} characterise non-termination
of a transition relation $T$ by the existence of an \emph{(open) recurrence
set}, i.e.  a nonempty set of states $N$ such that for each $s \in N$ there
exists a transition to some $s' \in N$.
The notion of open recurrence set is encoded by formula {\bf [ONT]} (Definition~\ref{def:ont}).
% Dual to the termination problem, when formulating $L$'s non-termination, we only need to consider an under-approximation of its transition
% relation~$T_u$. 
% Again, this distinction is mostly important when dealing with nested loops,
% as illustrated in Section~\ref{sec:env}.

\begin{theorem}
\label{thm:ont}
 The loop $L(G, T)$ has an infinite execution iff formula {\bf [ONT]} (Definition~\ref{def:ont}) is satisfiable.
\end{theorem}

@@

If this formula is satisfiable, $N$ is an open recurrence set for $L$, which proves
$L$'s non-termination. The issue with this formula is the additional level of quantifier alternation as compared to the synthesis fragment
(it is an $\exists \forall \exists$ formula).  To eliminate the innermost existential quantifier,
we introduce a Skolem function $S$, which we then existentially quantify over.  This results in
formula {\bf [SNT]} (Definition~\ref{def:snt}).

This extra second-order term introduces some complexity to the formula, which
we can avoid if the transition relation $T$ is deterministic.
\begin{definition}[Determinism]
 A relation $T$ is deterministic iff each state $x$ has exactly one successor under $T$:
 $$\forall x . \exists x' . T(x, x') \wedge \forall x'' . T(x, x'') \rightarrow x'' = x'$$
\end{definition}
In order to describe a deterministic \emph{program} in a way that still allows us
to sensibly talk about termination, we assume the existence of a special sink
state $s$ with no outgoing transitions and such that $\lnot G(s)$ for any
of the loop guards $G$.  The program is deterministic if its transition
relation is determinstic for all states except $s$.

When analysing a deterministic loop, we can make use of the notion of a \emph{closed recurrence set}
introduced by Chen et al.~in~\cite{DBLP:conf/tacas/ChenCFNO14}:  for each
state in the recurrence set $N$, \emph{all} of its successors must be in $N$.
The existence of a closed recurrence set is equivalent to the satisfiability
of formula {\bf [CNT]} in Definition~\ref{def:cnt}, which is already in the synthesis
fragment without needing Skolemization.

@@

@@

It is clear that this loop has many non-terminating executions,
e.g. the execution where \lstinline!nondet()! always returns 0.
However it is equally clear that each state has a successor
that exits the loop, i.e. when \lstinline|nondet()| returns
the value currently stored in \lstinline|x|.  So this loop
has an open recurrence set, but no closed recurrence set
and hence we cannot give a proof of its non-termination
with formula {\bf [CNT]} and instead must use formula {\bf [SNT]}.

@@

\subsection{An Isolated, Nested Loop}
\noindent {\bf Termination.} If a loop $L(G, T)$ has another loop $L'(G', T')$ nested inside it, we cannot directly use {\bf [UT]}
to express the termination of $L$.  This is because the single-step transition relation $T$ must
include the transitive closure of the inner loop $T'^*$, and we do not have a transitive closure
operator in our logic.  Therefore to encode the termination of $L$, we construct an overapproximation
$T_o \supseteq T$ and use this in formula {\bf [UT]} to specify a ranking function.
Rather than explicitly construct $T_o$ using, for example, abstract interpretation, we add constraints to
our formula which encode the fact that $T_o$ is an overapproximation of $T$, and that it is
precise enough to show that $R$ is a ranking function.

% All we need to do in order to use {\bf [CT]} in any context %environment and regardless of the complexity of the loops's body 
% is generate constraints over-approximating %abstracting 
% the environment and the loop's body. %: for termination, we need over-approximating abstractions, whereas for non-termination 
As the generation of such constraints is standard and covered by several other works \cite{DBLP:conf/pldi/GrebenshchikovLPR12,DBLP:conf/pldi/GulwaniSV08}, 
we will not provide the full algorithm, but rather illustrate it through the example in Figure~\ref{fig:environment-model}.
Our technique proves the termination of each loop separately and in program order, 
with the exception of nested loops, whose termination we encode in the same formula.
Following the same design, let us assume that we have already proven the termination of loop $L_1$, and we are now trying to prove that $L_2$ and $L_3$ terminate, which we encode
as the conjunction of constraints on the right side of the figure:
%% =======
%% %
%% As the generation of such constraints is standard and covered by several
%% other
%% works~\cite{DBLP:conf/pldi/GrebenshchikovLPR12,DBLP:conf/pldi/GulwaniSV08},
%% we will not provide the full algorithm, but rather illustrate it through the
%% example in Figure~\ref{fig:environment-model}.  Let us assume that we have
%% already proven the termination of loops $L_1$ and $L_3$, and we are now
%% trying to prove that $L_2$ terminates, which we encode as the conjunction of
%% constraints on the right-hand side of the figure:
%% %
%% >>>>>>> ee2c21bbc574ab36e7d6739c0e82c7a03d930789
\begin{itemize}
\item $W_1$ and $W_2$ are (over-approximating) inductive invariants for $L_1$ and $L_2$, respectively.
\item $S$ is a summary of $L_3$ that over-approximates the transitive closure of its transition relation.
\item If the formula is satisfiable, then $R_2$ and $R_3$ are ranking functions for $L_2$ and $L_3$, respectively.
\end{itemize}

@@

\noindent {\bf Non-Termination.}
Dually to termination, when proving non-termination, we need to
under-approximate the loop's body and apply formula {\bf [CNT]}.
%
% Recall that there is one more constraint that we must obey when doing so: we
% must stay inside the synthesis fragment!  This becomes more challenging now
% then when dealing with over-approximations due to the existential nature of
% under-approximations: they require an additional existential quantifier,
% which might result in an additional quantifier alternation, i.e.~$\exists
% \forall \exists$.
% 
% This issue does not arise when abstracting the environment, as its
% under-approximation and the loop under consideration make use of existential
% quantification at the same nestedness level.
%
%
Under-approximating the inner loop can be done with a nested existential quantifier, resulting in
$\exists \forall \exists$ alternation which we could eliminate with skolemization.  
However, we observe that
%As we try to avoid an extra second order unknown, 
unlike a ranking function,
the defining property of a recurrence set is \emph{non relational} -- if we
end up in the recurrence set, we do not care exactly where we came from as
long as we know that it was also somewhere in the recurrence set.  
This allows us to cast non-termination of nested loops as the formula shown in
Figure~\ref{fig:nonterm-nested}, which does not use a Skolem function.

@@

\noindent {\bf Conditional termination.}
Given a loop $L(G,T)$, if $L$'s termination depends on the state it begins
executing in, we say that $L$ is \emph{conditionally terminating}.
The information we require of the environment is a predicate $I$ which
over-approximates the set of states that $L$ may begin executing in.
That is to say, for each state $x$ that is reachable on entry to $L$,
we have $I(x)$.

\begin{theorem}
\label{thm:ct}
 The loop $L(G, T)$ terminates when started in any state satisfying $I(x)$ iff formula {\bf [CT]}
 (Defintion~\ref{def:ct}) is satisfiable.
\end{theorem}
% 
% \begin{proof}
%  Do this proof.
% \end{proof}

If formula {\bf [CT]} is satisfiable, two witnesses are returned:
\begin{itemize}
\item $W$ is an inductive invariant of $L$ that is established by the initial states $I$ if the loop
guard $G$ is met.
\item $R$ is a ranking function for $L$ as restricted by $W$ -- that is to say, $R$ need only
be well founded on those states satisfying $W \wedge G$.  Since $W$ is an inductive invariant of $L$,
$R$ is strong enough to show that $L$ terminates from any of its initial states.
\end{itemize}

However, formula {\bf [CT]} is satisfiable, as witnessed by:
\begin{align*}
R(\langle x,y\rangle) & = x\\
W(\langle x, y \rangle ) & \triangleq y  = 1
\end{align*}

This constitutes a proof that the program as a whole terminates, since the loop always begins
executing in a state that guarantees its termination.\\

% \noindent {\bf Over-approximating the environment for termination.}
%% While proving that loop $L$ terminates, we create constraints on the environment in the
%% form of the predicate $I$ and the over-approximating summary relation $T_o$:
%% In the presence of nested loops, 
%% for each loop $L'$ nested in the body
%% of $L$, we must generate constraints saying that $T_o$ over-approximates $L'$, i.e. it is an over-approximation of the transitive closure of $L'$'s transition relation.
%% Similarly, for each loop $L''$ appearing before $L$ in the program's control flow, we
%% must add constraints saying that $I$ over-approximates the states reachable via $L''$.
%The details of how to create these constraints are standard, so
%rather than presenting the algorithm for generating the constraints, we direct the
%reader's attention to Figure~\ref{fig:environment-model}.  
%
% %\noindent {\bf Environment abstractions.}
% For a loop $L=(G,T)$, the formulae {\bf [CT]} %and {\bf [CNT]} 
% from Definition~\ref{def:ct} 
% %and Definition~\ref{def:cnt} 
% is in terms of 
% $T_o$ and $I$, which over-approximate the transition relation~$T$ and the set of states that $L$ may begin executing in, respectively. 
% The reason for using over-approximations [instead of exact relations] %when formulating the (non-)termination of an isolated loop 
% is so that we can use the same formulae regardless of the environment the loop is executing in, or 
% the complexity of its body, e.g.~if it contains nested loops.
% 
% \noindent {\bf Under-approximating the environment for non-termination.}

@@

\subsection{Generalised Termination and Non-termination Formula}
At this point, we know how to construct two formulae for a loop $L$: one that is
satisfiable iff $L$ is terminating and another that is
satisfiable iff it is non-terminating.  We will call these formulae $\phi$ and $\psi$
respectively:
\begin{align*}
 \exists P_N . \forall x, x' . \phi(W_N, x, x') \\
 \exists P_T . \forall x, x' . \psi(W_T, x, x')
\end{align*}
We can combine these:
\begin{align*}
 (\exists P_N . \forall x, x'. \phi(W_N, x, x')) \vee (\exists P_T . \forall x, x', \psi(W_T, x, x'))
\end{align*}
Which simplifies to:
\begin{definition}[Generalised Termination Formula]
\begin{align*}
 \exists P_N, P_T. \forall x, x', y, y', \phi(P_N, x, x') \vee \psi(P_T, y, y') 
\end{align*}
\end{definition}

Since $L$ either terminates or does not terminate, this formula is a tautology in the synthesis fragment.
A solution to the formula would include witnesses $P_N$ and $P_T$, which are proofs of non-termination
and termination respectively.  Exactly one of these proofs will be valid, so we can check
first one and then the other.

% 
% Given a program, our algorithm attempts to prove termination or non-termination for each loop individually 
% by constructing a generalised formula for each of them.
% For nested loops, we generate one single formula for all the nested loops (as illustrated for termination in Figure~\ref{fig:environment-model}).
% 
% For an isolated loop $L(G,T)$, the generalised termination and non-termination formula is the following:
% %
%  \begin{align*}
%   \exists R, N, x_0 . \forall x, x',y,y' . & N(x_0) ~ \wedge \\ & N(x) \rightarrow G(x) ~ \wedge \\
% 							& N(x) \wedge T_u(x, x') \rightarrow N(x')\\
%   & \vee G(y) \wedge T_o(y, y') \rightarrow  R(y){>}0 \wedge R(y){>}R(y')
%  \end{align*}
% %
% %
% where $T_u$ and $T_o$ under- and over-approximate the transition relation $T$, respectively.

@@

To evaluate our algorithm, we implemented two tools: one to generate a
termination specification from a C program, and one to synthesise a proof
from the specification.  We ran the resulting termination prover on 47
benchmarks taken from the literature and SV-COMP'15~\cite{svcomp}.  To
provide a comparison point, we also ran {\sc ARMC}~\cite{armc-website} on
the same benchmarks.  Each tool was given a time limit of $180\,s$, and was
run on an unloaded 8-core 3.07\,GHz Xeon X5667 with 50\,GB of RAM.  The
results of these experiments are shown in Figure~\ref{fig:experiments}.
